### 2023-09-01 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d9e5089ca1aeb57d00b06ed661d22af1941cf001 🚀

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{20} and filtered it to ``known'' hosts
(established before the emergence of SARS-CoV-2) and ``novel'' hosts
(confirmed through sampling and competence assays since the initial data
collection). The original database was assembled by a combination of
data mining and literature surveys, including automated alerts on the
``bats'' and ``coronavirus'' keywords to identify novel empirical
evidence of bats-betacoronaviruses associations; this yielded a total of
126 known hosts, 47 of which were novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{56} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
100kmx100km at the equator. For every pixel in the resulting raster
where at least one bat host of \emph{Betacoronavirus} was present, we
extract the species pool (list of all known bat hosts), which was used
to calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{58} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{59} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{60} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{61} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{62} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{63} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{64} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{65} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{66} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
with ModelFinder\textsuperscript{69} ultrafast bootstrap
approximation\textsuperscript{70} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{71} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{72,73} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
\CSLLeftMargin{1. }
\CSLRightInline{Plowright, R. K. \emph{et al.} Pathways to zoonotic
spillover. \emph{Nature Reviews Microbiology} \textbf{15}, 502--510
(2017).}

\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
\CSLLeftMargin{2. }
\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}

\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
\CSLLeftMargin{3. }
\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
\textbf{20}, 299--314 (2022).}

\leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
spillover risk of bat SARS-related coronaviruses in southeast asia.
\emph{Nature communications} \textbf{13}, 1--12 (2022).}

\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
can be generalists: Resolving the {`parasite paradox'} and implications
for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
151--162 (2010).}

\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
(2004).}

\leavevmode\hypertarget{ref-Geoghegan2017ComAna}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
Comparative analysis estimates the relative frequencies of co-divergence
and cross-species transmission within viral families. \emph{PLOS
Pathogens} \textbf{13}, e1006215 (2017).}

\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
\CSLLeftMargin{8. }
\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
Coevolution}. (University Of Chicago Press, 2005).}

\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
\CSLLeftMargin{9. }
\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
(University of Chicago Press, 1994).}

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
\CSLLeftMargin{10. }
\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
\textbf{34}, 611--612 (1980).}

\leavevmode\hypertarget{ref-Price2002MacThe}{}%
\CSLLeftMargin{11. }
\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
Macroecological Patterns}. (Cambridge University Press, 2002).
doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
\CSLLeftMargin{12. }
\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
the geographic mosaic theory of coevolution. \emph{Heredity}
\textbf{98}, 249--258 (2007).}

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
\CSLLeftMargin{13. }
\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
J. D. Coevolution of symbiotic mutualists and parasites in a community
context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
(2007).}

\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
\CSLLeftMargin{14. }
\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
Server for Biology} 2020.05.31.116061 (2020)
doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}

\leavevmode\hypertarget{ref-Anthony2013CorBat}{}%
\CSLLeftMargin{15. }
\CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
Mexico. \emph{The Journal of General Virology} \textbf{94}, 1028--1038
(2013).}

\leavevmode\hypertarget{ref-Goes2013NovBat}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
Brazil and Mexico. \emph{Emerging Infectious Diseases} \textbf{19},
1711--1713 (2013).}

\leavevmode\hypertarget{ref-Goes2016GenDiv}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
coronaviruses in the Atlantic Forest hotspot biome, Brazil.
\emph{Infection, Genetics and Evolution: Journal of Molecular
Epidemiology and Evolutionary Genetics in Infectious Diseases}
\textbf{44}, 510--513 (2016).}

\leavevmode\hypertarget{ref-Brandao2008CorDet}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
the vampire bat Desmodus rotundus. \emph{Brazilian Journal of Infectious
Diseases} \textbf{12}, 466--468 (2008).}

\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
646--650 (2017).}

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe} (2022)
doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
molecular phylogenetic insights into the evolution of free-tailed bats
in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
Mammalogy} \textbf{93}, 12--28 (2012).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

\leavevmode\hypertarget{ref-Hu2017DisRic}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
bat SARS-related coronaviruses provides new insights into the origin of
SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, Eastern Africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
synthesis of ecological and evolutionary determinants of bat diversity
across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}

\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
World: A taxonomic and geographic database. \url{https://batnames.org/}
(2020).}

\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
\textbf{2013}, e187415 (2013).}

\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
\CSLLeftMargin{41. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}

\leavevmode\hypertarget{ref-Caraballo2022CroTra}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
Coronaviruses in the Americas: Contrasting Patterns between
Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
\textbf{0}, e01411--22 (2022).}

\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

>>>>

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{\textbf{Becker2022OptPre?}} and filtered it
to ``known'' hosts (established before the emergence of SARS-CoV-2) and
``novel'' hosts (confirmed through sampling and competence assays since
the initial data collection). The original database was assembled by a
combination of data mining and literature surveys, including automated
alerts on the ``bats'' and ``coronavirus'' keywords to identify novel
empirical evidence of bats-betacoronaviruses associations; this yielded
a total of 126 known hosts, 47 of which were novel hosts. This
host--virus list of interactions was obtained through a comprehensive
aggregation of GenBank data as well as systematic literature
searches,\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
such that we have high confidence in its fitness for the purpose of
inference at a large spatial scale.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{\textbf{IUCN2021IucRed?}} The IUCN data have been
assembled to support wildlife conservation efforts, and therefore we do
not expect that they are biased by wildlife disease sampling efforts or
priority. The range maps were subsequently rasterized using the
\texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{\textbf{RouaultEven2022GdaOgr?}} at a
resolution of approximately 100kmx100km at the equator. For every pixel
in the resulting raster where at least one bat host of
\emph{Betacoronavirus} was present, we extract the species pool (list of
all known bat hosts), which was used to calculate the following risk
assessment components: bat phylogenetic diversity, bat compositional
uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{\textbf{Faith1992ConEva?}} based on a recent
synthetic tree with robust time calibration, covering about 6000
mammalian species.\textsuperscript{\textbf{Upham2019InfMam?}} Faith's PD
measures the sum of unique branches from an arbitrary root to a set of
tips, and comparatively larger values indicate a more phylogenetic
diverse species pool. We measured phylogenetic diversity starting from
the root of the entire tree (and not from Chiroptera); this bears no
consequences on the resulting values, since all branches leading up to
Chiroptera are only counted once per species pool, and (as we explain
when describing the assembly of the composite risk map), all individual
risk components are ranged in {[}0,1{]}. This measure incorporates a
richness component, which we chose not to correct for; the
interpretation of the phylogenetic diversity is therefore a weighted
species richness, that accounts for phylogenetic over/under-dispersal in
some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{\textbf{Legendre2013BetDiv?}} LCBD works
from a species-data matrix (traditionally noted as \(\mathbf{Y}\)),
where species are rows and sites are columns, and a value of 1 indicates
occurrence. We extracted the Y matrix assuming that every pixel
represents a unique location, and following best
practices\textsuperscript{\textbf{Legendre2019SpaTem?}} transformed it
using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results
by\textsuperscript{\textbf{Dansereau2022EvaEco?}} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{3} especially for
beta-coronaviruses.\textsuperscript{4} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{\textbf{Seekell2018GeoLak?}} In this approach,
every risk component gets assigned a component in the RGB color model
(phylogenetic diversity is green, compositional uniqueness is red, and
viral sharing is blue). In order to achieve a valid RGB measure, all
components are re-scaled to the {[}0,1{]} interval, so that a pixel with
no sharing, no phylogenetic diversity, and no compositional uniqueness
is black, and a pixel with maximal values for each is white. This
additive model conveys both the intensity of the overall risk, but also
the nature of the risk as colors diverge towards combinations of values
for three risk components. Out of the possible combinations, the most
risky in terms or rapid diversification and spillover potential is high
phylogenetic diversity and low viral
sharing,\textsuperscript{\textbf{Gomulkiewicz2000HotSpo?}} in that this
allows multiple independent host-virus coevolutionary dynamics to take
place in the same location. In the colorimetric space, this correspond
to yellow -- because the HSV space is more amenable to calculations for
feature extraction,\textsuperscript{\textbf{Keke2010StuSki?}} we
measured the risk level by calculating the angular distance of the hue
of each pixel to a reference value of 60 (yellow), and weighted this
risk level by the value component. Specifically, given a pixel with
colorimetric coordinates \((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{\textbf{Katoh2013MafMul?}} v1.4.0 (Algorithm
FFT-NS-2, Scoring matrix 200PAM / k=2, gap open penalty 1.53m offset
value 0.123) and a maximum likelihood tree reconstructed in
IQ-TREE\textsuperscript{\textbf{Nguyen2015IqtFas?}} v1.6.12 with
ModelFinder\textsuperscript{\textbf{Kalyaanamoorthy2017ModFas?}}
ultrafast bootstrap
approximation\textsuperscript{\textbf{Hoang2018UfbImp?}} with a general
time reversible model with empirical base frequencies and the
5-discrete-rate-category FreeRaye model of nucleotide substitution
(GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{\textbf{Isaac2007MamEdg?}} for each of
the viruses in the tree, then averaged these at the bat species level,
projected these values onto their geographic distributions, and averaged
across every bat found in a given pixel. As such, this can be thought of
as a map of the mean evolutionary distinctiveness of the known viral
community believed to be associated with a particular subset of bats
present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method
from,\textsuperscript{\textbf{Kreft2007GloPat?},\textbf{Kreft2010FraDel?}}
who proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in
\texttt{ape};\textsuperscript{\textbf{Paradis2019ApeEnv?}} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

\hypertarget{data-availability-statement}{%
\subsection{Data availability
statement}\label{data-availability-statement}}

The code to reproduce these analyses, as well as the data (with the
exception of the IUCN rangemaps, which must be downloaded from their
website) are available in the
\href{https://github.com/viralemergence/betamap}{\texttt{viralemergence/betamap}}
repository on GitHub.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{1. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{2. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{3. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

\leavevmode\hypertarget{ref-Latinne2020Origin}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
Server for Biology} 2020.05.31.116061 (2020)
doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
